Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-06-11 article.tex
typos

\begin{corollary}
\label{corollary:max is ascent}
Let $\pi \in \in \AV_n(213,231)$ and $1 \leq i < n$.
Then,
(1) 
if $\pi[i]$ is an ascent element then $\pi[i]$ is the minimum element 
of $\pi[i:]$
and 
(2)
if if $\pi[i]$ is a descent element then $\pi[i]$ is the maximum element 
of $\pi[i:]$
\end{corollary}

@@

\begin{proof}
  The forward direction is obvious.
  We prove the backward direction by induction on the size of the motif
  $\sigma$.
  The base case is a motif of size $2$.
  Suppose first that $\sigma = 12$ and hence $s_\sigma = a$.
  Let $t = \pi_{i_1}\pi_{i_2}$, $i_1 < i_2$, be a subsequence of $\pi$
  such that $s_t$ occurs as a subsequence in $s_\pi$.
  Since $s_\sigma = s_t = a$, this reduces to saying that
  $\pi_{i_1} < \pi_{i_2}$, and hence $t$ is an occurrence of $\sigma = 12$ in $\pi$.
  A similar argument shows that the lemma holds true for $\sigma = 21$.
  Now, assume that the lemma is true for all motifs up to size $k \geq 2$.
  Let $\sigma \in \Av{k+1}{231,213}$ and
  let $t = \pi_{i_1}\pi_{i_2}\,\ldots\,\pi_{i_{k+1}}$,
  $i_1 < i_2 < \ldots < i_{k+1}$,
  be a subsequence of $\pi$ of length $k+1$ such that
  $s_t = s_\sigma$.
  Suppose that $s_t$ occurs in $s_\pi$ as a subsequence.
  Then, $s_t[2:] = s_\sigma[2:]$ occurs as a subsequence in $s_\pi[2:]$.
  But $\sigma[2:] \in \Av{k}{231,213}$ and hence,
  by the inductive hypothesis, it follows that
  $t[2:]$ is an occurrence of $\sigma[2:]$ in $\pi[2:]$.
  Moreover, if $s_\sigma[1] = a$ (resp. $s_\sigma[1] = d$)
  then $\sigma[1]$ is the minimum (resp. maximum) element of $\sigma$
  (Lemma~\ref{lemma:first element is 1 or n}),
  and hence, since $s_t = s_\sigma$ occurs as a subsequence in $s_\pi$,
  $t[1]$ is the minimum (resp. maximum) element of $t$.
  Therefore, $t$ is an occurrence of $\sigma$ in $\pi$.
  \qed
\end{proof}

@@

This section focuses on the pattern containment problem
in case only the pattern $\sigma$ avoids $231$ and $213$.
We need to consider a specific decomposition $\factora$ of $\sigma$ into factors.
For $\sigma \in S_k$, the \emph{factorization}
$\sigma = \factor{\pmotif}{p} \factor{\pmotif}{p-1} \ldots \factor{\pmotif}{1}$
is defined as follows:
(i) $\sigma[k]$ is the last letter of $\factor{\pmotif}{1}$,
(ii)for every $1 \leq i < k$,
if $s_\sigma[i] = s_\sigma[i+1]$ then
$\sigma[i]\sigma[i+1]$ is part of the same factor.
For every factor $\factor{\pmotif}{j}$ of the decomposition, we let
$\first{\pmotif}{j}$ stand for the leftmost letter of $\factor{\pmotif}{j}$,
and $\firsti{\pmotif}{j}$ stand for the index in $\sigma$
of the leftmost letter of $\factor{\pmotif}{j}$.
For example,
for $\sigma = 981237654$,
we have
$s_\sigma = ddaaaddd$ and hence
$\sigma = \factor{\pmotif}{3} \factor{\pmotif}{2} \factor{\pmotif}{1}$ with
$\factor{\pmotif}{1} =7654$, $\factor{\pmotif}{2} = 123$ and $\factor{\pmotif}{3} = 98$.
Furthermore,
$\first{\pmotif}{1} = 7$, $\first{\pmotif}{2} = 1$, $\first{\pmotif}{3} = 9$,
$\firsti{\pmotif}{1} = 6$, $\firsti{\pmotif}{2} = 3$ and $\firsti{\pmotif}{3} = 1$.

@@

\begin{lemma}
\label{lemma:whereIsMax}
Given a permutation $\sigma \in \AV(213,231)$ and a suffix of its decomposition $\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$, if
$\firsti{\sigma}{i}$ is an ascent element (respectively  if $\firsti{\sigma}{i}$ is a descent element) then the maximal (resp. minimal) element of $\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$ is $\first{\sigma}{i-1}$.
\end{lemma}

@@

For easy readability, we now define the set
$\SET{\pmotif}{\ptext}{i}{j}$ as follows:
if $s \in \SET{\pmotif}{\ptext}{i}{j}$ then
$s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$
and the element matching $\first{\pmotif}{i}$ is $\ptext[j]$.

\begin{lemma}
Given a permutation $\pmotif$ and $\factor{\pmotif}{i}$ an ascent (respectively descent) factor.
If $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that the element matching $\first{\pmotif}{i-1}$ is minimal (resp. maximal), for all subsequence 
$s' \in \SET{\pmotif}{\ptext}{i}{j}$ and for all subsequence $t$ of $\ptext$, such that $t=t's'$, if $t$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ then the subsequence $t's$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$.
\end{lemma}

\begin{proof}
By definition $s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$. To prove that $ts$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ we prove that every element of t is larger than every element of $s$. If $ts'$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ so every element of $t$ is larger than every element of $s'$. Moreover the maximal element of $s$ is smaller than  the maximal element of $s'$ so every element of $s$ is smaller than every element of $s'$ thus every element of $s$ is smaller than every element of $t$. We use a similar argument if $\factor{\pmotif}{i}$ a descent factor.
$\qed$
\end{proof}

\begin{corollary}
Given a permutation $\pmotif$ and $\factor{\pmotif}{i}$ an ascent (respectively descent) factor.
If $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). There is a match of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ matched to $\ptext[j]$ if and only if there exists $t$ such that $ts$ is a match of $\pmotif$ in $\ptext$. 
\end{corollary}

This corollary gives one a safe way to choose a match. Indeed if one does not find a match such that a well chosen subsequence is a suffix of the motif then there is no match at all.

@@

Consider the following problem :\\
$
\LM{\pmotif}{\ptext}{i}{j} =
\begin{cases}

@@

	\text{The minimal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & \text{ a descent element } \\	
	\text{ and that maximize the minimal value} & \\
	\text{ and $\first{\sigma}{i}$ is matched with $\ptext[j]$} & \\
\end{cases}
$

This problem can be solve by induction.\\

\textbf{BASE :} \\
$
\LM{\pmotif}{\ptext}{1}{j} =
\begin{cases}
		\min_{j<j'}\{\infty\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LIS(j,j',\ptext[j'])  $}\}&\text{ an ascent element } \\
		&\\
		\max_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LDS(j,j',\ptext[j']) $}\}&\text{ a descent element }\\
\end{cases}
$

\textbf{STEP :} \\
$
\LM{\pmotif}{\ptext}{i}{j} =
\begin{cases}
	\min \{\infty\} \cup  \AF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is an ascent element}\\
	\max \{0\} \cup  \DF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is  a descent element}\\
\end{cases}
$

%
%With $\AF{\pmotif}{\ptext}{i}{j}$ ($\DF{\pmotif}{\ptext}{i}{j}$) the set of element such that
%if  $\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$ then there exists
%a match of $\pmotif[\firsti{\pmotif}{i-1}:]$ in $\ptext[k:]$ and
%the element $\first{\pmotif}{i-1}$ is matched with element $\ptext[k]$ and
%there is a match of $\factor{\pmotif}{i}$ in $\ptext[j:k-1]$ and
%the element $\first{\pmotif}{i}$ is matched with element $\ptext[j]$ and every element is inferior (superior) to the match of
%$\pmotif[\firsti{\pmotif}{i-1}:]$ in $\ptext[k:]$.

With $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ the set of element such that if  $\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$ or if $\ptext[k] \in \DF{\pmotif}{\ptext}{i}{j}$ then there exists a match of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$ and the element $\first{\pmotif}{i-1}$ is matched with element $\ptext[k]$. Given that one has the set $\SET{\pmotif}{\ptext}{i-1}{k}$, to decide if $\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$, one needs to find an occurrence $t$ of $\factor{\pmotif}{i}$ in $\ptext[j:k-1]$ such there exists $s \in \SET{\pmotif}{\ptext}{i-1}{k}$ and $ts$ is a match of $\pmotif[\firsti{\pmotif}{i}:]$. 
Formally we define $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ such as:\\

$\AF{\pmotif}{\ptext}{i}{j}=\{\ptext[j']|\text{$j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq 0$}$ \\
\indent \indent \indent and  $|\factor{\pmotif}{i}| \leq LIS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$ \\

$\DF{\pmotif}{\ptext}{i}{j}=\{\ptext[j']|\text{$j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq \infty$}$ \\
\indent \indent \indent and  $|\factor{\pmotif}{i}| \leq LDS_{\ptext}(j,j'-1,\first{\pmotif}{i-1}) \}$ \\

With $LIS_{\ptext}(i,j,k)$ (respectively $LDS_{\ptext}(i,j,k)$) is the longest increasing (resp. decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
smaller (resp. bigger) than $k$.
$LIS_{\ptext}$ and $LDS_{\ptext}$ can be computed in $O(|\ptext|^2\log(\log(|\ptext|)))$ (see \cite{Bespamyatnikh00enumeratinglongest}).\\

For the base case,
one is looking for a match for the first factor.
If the factor is an ascent (respectively descent) factor, then
one has to find an increasing (resp. decreasing) subsequence
in the text of same size or longest that
the size of the first factor.
To find the "safest" solution
one must assure that the right most element
of the sequence is minimal (resp. maximal).\\
For the induction, it follows the same idea
except that one must assure that
every element of the subsequence is larger (resp. smaller)
than the maximum (resp. minimal) element of the rest of the match which is given by the inductive problem of
the previous factor.\\

There exists a match of $\pmotif$ in $\ptext$ if and only if
there exists a $LM(n,i)\neq 0$ and $LM(n,i)\neq \infty$ for $1 \leq i \leq |\ptext|$,
with $n$ the number of factor in $\pmotif$.
Moreover the basic case can be computed in $O(n^2)$
and the induction on $O(kn^2)$ time and space.
$\qed$
\end{proof}

@@

		This section is devoted to presenting a polynomial-time algorithm for deciding whether
	a $(213,231)$-avoiding bivincular pattern occurs in a permutation.

We introduce property of bivincular permutation for a motif avoiding both $213$ and $231$.

\begin{lemma}
\label{lemma:ascentDescentAscent}
If $\overline{\sigma[i]\sigma[j]}$ (thus $\sigma[j]=\sigma[i]+1$) and if $\sigma[i]$ is an ascent (respectively decent) element and $\sigma[i]+1$ is an ascent (resp. decent) element then $i<j$ ($j>i$) and for every $k$, $i<k<j$ ($j>k>i$), $\pi[k]$ is a descent (resp. ascent) element.    
\end{lemma}

\begin{proof}
Suppose that there exists $k$, $i<k<j$,
such that $\pmotif[k]$ is ascent. Ascent elements are increasing  so $\sigma[i]<\sigma[k]<\sigma[j]$ which is in contradiction that with $\sigma[j]=\sigma[i]+1$.
We use a similar argument if $\sigma[i]$ is a descent element
$\qed$
\end{proof}

@@

\begin{proof}
We consider the following problem :\\
Given a bivincular motif $\pbmotif$ with $\pmotif \in \Av{n_\pmotif}{231,213} $, and a text $\ptext \in \Av{n_\ptext}{231,213}$.
$$
\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
\begin{cases}
	true 	& \text{If $\pbmotif[i:]$ occurs in $\ptext[j:]$}\\
			& \text{with every element is in $[\lb,\ub]$}\\
			& \text{and element $\pmotif[i]$ occurs in $\ptext[j]$ }\\
	false 	& otherwise\\
\end{cases}
$$

$\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ is closed under induction.
It can be solved by means of the following relations:\\
\textbf{BASE :} \\
$$
\PM{\pbmotif}{\ptext}{\lb}{\ub}{n_\pmotif}{j}=
\begin{cases}
	true 	& \text{if $\ptext[j] \in [\lb,\ub ]$}\\
			& \text{and if ${\pmotif[n_\pmotif]}_\lrcorner$ then $j=n_\ptext$}\\
			& \text{and if ${\pmotif[n_\pmotif]}^\urcorner$ then $\ptext[j]=\ub=n_\pmotif$}\\
			& \text{and if  $^\ulcorner{\pmotif[n_\pmotif]}$ then $\ptext[j]=\lb=1$ } \\
			& \text{and if  $\overline{(\pmotif[n_\pmotif]-1)\pmotif[n_\pmotif] }$ then $\ptext[j]=\lb$ }  \\
			& \text{and if  $\overline{(\pmotif[n_\pmotif]\pmotif[n_\pmotif]-1)}$ then $\ptext[j]=\ub$}  \\

@@

\textbf{STEP :} \\
For clarity of reading we consider 3 cases for the problem $\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ :
\begin{itemize}
	\item If $\ptext[j] \notin [\lb,\ub]$ then :
	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j} = false
	$$

@@

			the inductive problems after $\PM{\pbmotif}{\ptext}{*}{*}{i}{*}$ have for lower bound $\ptext[j]+1$. $\PM{\pbmotif}{\ptext}{*}{*}{i'}{*}$ is either an inductive case or a base case, in either way  $\PM{\pbmotif}{\ptext}{*}{*}{i'}{*}$ is true if and only if the occurrence of $\sigma[i']$ is equal to lower bound which is equal to $\sigma[i']+1$ so the occurrence of $\pmotif[i]$ and $\pmotif[i']$ are consecutive in value.
		\end{itemize}

\end{itemize}
$\qed$
\end{proof}

%\begin{proof}
%We consider the following problem :\\
%Given a bivincular motif $\pbmotif$ with $\pmotif \in \Av{n_\pmotif}{231,213} $, and a text $\ptext \in \Av{n_\ptext}{231,213}$.
%$$
%\PM{\pbmotif}{\ptext}{\min}{\max}{i}{j}=
%\begin{cases}
%	true 	& \text{If $\pbmotif[i:]$ is a motif of $\ptext[j:]$}\\
%			  & \text{with every element in $]\min,\max[$}\\
%	false 	& otherwise\\
%\end{cases}
%$$
%$\PM{\pbmotif}{\ptext}{\min}{\max}{i}{j}$ is closed under induction.
%It can be solved by means of the following relations:\\
%
%\textbf{BASE :} \\
%
%For the base case, one matches the right most element of the motif. If the motif has a right bottom corner (if $n_\pmotif \in  \x$) then one must assure that the right most element of the motif is matched to the right most of the text. If the motif has a right top corner (if $n_\pmotif \in  \y$) and the right most element of the motif is maximal (minimal) then one must assure that right most element of the motif is matched to the maximal (minimal) element of the text. If there is not corner, there is no condition on the match of the right most element.
%
%if $n_\pmotif \notin  \x $ and $n_\pmotif \notin  \y$ : \\
%$
%\PM{\pbmotif}{\ptext}{\min}{\max}{n_\pmotif}{j}=
%\begin{cases}
%	true 	& \text{if $\min<\ptext[j]<\max$}\\
%	false 	& otherwise\\
%\end{cases}
%$
%%$\PM{\pbmotif}{\ptext}{\min}{\max}{n_\pmotif+1}{j}= true $
%\\
%
%if $n_\pmotif \in  \x $ and $n_\pmotif \notin  \y$ : \\
%$
%\PM{\pbmotif}{\ptext}{\min}{\max}{n_\pmotif}{j} =
%\begin{cases}
%	true 	& \text{if $\min<\ptext[j]<\max$}\\
%			& \text{and } j=n_\ptext\\
%	false 	& otherwise\\
%\end{cases}
%$
%%$\mathbb{P}_{\pbmotif,\min,\max}^{\ptext}(n_\pmotif+1,n_\ptext)= true $
%\\
%
%if $n_\pmotif \notin  \x$  and $n_\pmotif \in  \y$ and $\pmotif[n_\pmotif]$ is the maximal element: \\
%$
%\PM{\pbmotif}{\ptext}{\min}{\max}{n_\pmotif}{j}=
%\begin{cases}
%	true 	& \text{if $\min<\ptext[j]<\max$ }\\
%			& \text{and $\ptext[j]$ is the maximal element}\\
%	false 	& otherwise\\
%\end{cases}
%$
%%$\mathbb{P}_{\pbmotif,\min,\max}^{\ptext}(n_\pmotif+1,j)= true $\\
%
%if $n_\pmotif \in  \x$  and $n_\pmotif \in  \y$ and $\pmotif[n_\pmotif]$ is the maximal element: \\
%$
%\PM{\pbmotif}{\ptext}{\min}{\max}{n_\pmotif}{j}=
%\begin{cases}
%	true 	& \text{if $\min<\ptext[j]<\max$}\\
%			& \text{and } j=n_\ptext\\
%			& \text{and $\ptext[n_\ptext]$ is the maximal element}\\
%	false 	& otherwise\\
%\end{cases}
%$
%%$\mathbb{P}_{\pbmotif,\min,\max}^{\ptext}(n_\pmotif+1,n_\ptext)= true $\\
%
%if $n_\pmotif \notin  \x$  and $n_\pmotif \in  \y$ and $\pmotif[n_\pmotif]$ is the minimal element: \\
%$
%\PM{\pbmotif}{\ptext}{\min}{\max}{n_\pmotif}{j}=
%\begin{cases}

>>>>

\begin{corollary}
\label{corollary:max is ascent}
Let $\pi \in \in \AV_n(213,231)$ and $1 \leq i < n$.
Then,
(1) 
if $\pi[i]$ is an ascent element then $\pi[i]$ is the minimal element 
of $\pi[i:]$
and 
(2)
if if $\pi[i]$ is a descent element then $\pi[i]$ is the maximal element 
of $\pi[i:]$
\end{corollary}

@@

\begin{proof}
  The forward direction is obvious.
  We prove the backward direction by induction on the size of the motif
  $\sigma$.
  The base case is a motif of size $2$.
  Suppose first that $\sigma = 12$ and hence $s_\sigma = a$.
  Let $t = \pi_{i_1}\pi_{i_2}$, $i_1 < i_2$, be a subsequence of $\pi$
  such that $s_t$ occurs as a subsequence in $s_\pi$.
  Since $s_\sigma = s_t = a$, this reduces to saying that
  $\pi_{i_1} < \pi_{i_2}$, and hence $t$ is an occurrence of $\sigma = 12$ in $\pi$.
  A similar argument shows that the lemma holds true for $\sigma = 21$.
  Now, assume that the lemma is true for all motifs up to size $k \geq 2$.
  Let $\sigma \in \Av{k+1}{231,213}$ and
  let $t = \pi_{i_1}\pi_{i_2}\,\ldots\,\pi_{i_{k+1}}$,
  $i_1 < i_2 < \ldots < i_{k+1}$,
  be a subsequence of $\pi$ of length $k+1$ such that
  $s_t = s_\sigma$.
  Suppose that $s_t$ occurs in $s_\pi$ as a subsequence.
  Then, $s_t[2:] = s_\sigma[2:]$ occurs as a subsequence in $s_\pi[2:]$.
  But $\sigma[2:] \in \Av{k}{231,213}$ and hence,
  by the inductive hypothesis, it follows that
  $t[2:]$ is an occurrence of $\sigma[2:]$ in $\pi[2:]$.
  Moreover, if $s_\sigma[1] = a$ (resp. $s_\sigma[1] = d$)
  then $\sigma[1]$ is the minimal (resp. maximal) element of $\sigma$
  (Lemma~\ref{lemma:first element is 1 or n}),
  and hence, since $s_t = s_\sigma$ occurs as a subsequence in $s_\pi$,
  $t[1]$ is the minimal (resp. maximal) element of $t$.
  Therefore, $t$ is an occurrence of $\sigma$ in $\pi$.
  \qed
\end{proof}

@@

This section focuses on the pattern containment problem
in case only the pattern $\sigma$ avoids $231$ and $213$.
We need to consider a specific decomposition $\factora$ of $\sigma$ into factors.
For $\sigma \in S_k$, the \emph{factorization}
$\sigma = \factor{\pmotif}{p} \factor{\pmotif}{p-1} \ldots \factor{\pmotif}{1}$
is defined as follows:
(i) $\sigma[k]$ is the last letter of $\factor{\pmotif}{1}$,
(ii) for every $1 \leq i < k$,
if $s_\sigma[i] = s_\sigma[i+1]$ then
$\sigma[i]\sigma[i+1]$ is part of the same factor.
For every factor $\factor{\pmotif}{j}$ of the decomposition, we let
$\first{\pmotif}{j}$ stand for the leftmost letter of $\factor{\pmotif}{j}$,
and $\firsti{\pmotif}{j}$ stand for the index in $\sigma$
of the leftmost letter of $\factor{\pmotif}{j}$.
For example,
for $\sigma = 981237654$,
we have
$s_\sigma = ddaaaddd$ and hence
$\sigma = \factor{\pmotif}{3} \factor{\pmotif}{2} \factor{\pmotif}{1}$ with
$\factor{\pmotif}{1} =7654$, $\factor{\pmotif}{2} = 123$ and $\factor{\pmotif}{3} = 98$.
Furthermore,
$\first{\pmotif}{1} = 7$, $\first{\pmotif}{2} = 1$, $\first{\pmotif}{3} = 9$,
$\firsti{\pmotif}{1} = 6$, $\firsti{\pmotif}{2} = 3$ and $\firsti{\pmotif}{3} = 1$.

@@

\begin{lemma}
\label{lemma:whereIsMax}
Given a permutation $\sigma \in \AV(213,231)$ and a suffix of its 
decomposition $\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$, 
if $\firsti{\sigma}{i}$ is an ascent (resp. descent) element then the maximal 
(resp. minimal) element of $\factor{\sigma}{i}\factor{\sigma}{i-1}\ldots\factor{\sigma}{1}$ is 
$\first{\sigma}{i-1}$.
\end{lemma}

@@

For the sake of readability, we now define the set
$\SET{\pmotif}{\ptext}{i}{j}$ as follows:
if $s \in \SET{\pmotif}{\ptext}{i}{j}$ then
$s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$
and the element matching $\first{\pmotif}{i}$ is $\ptext[j]$.

\begin{lemma}
Let $\pmotif$ be a permutation where 
$\factor{\pmotif}{i}$ is an ascent (resp. descent) factor.
If $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that the element matching $\first{\pmotif}{i-1}$ is minimal (resp. maximal), for all subsequence 
$s' \in \SET{\pmotif}{\ptext}{i}{j}$ and for all subsequence $t$ of $\ptext$, such that $t=t's'$, if $t$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ then the subsequence $t's$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$.
\end{lemma}

\begin{proof}
By definition $s$ is an occurrence of $\pmotif[\firsti{\pmotif}{i}:]$. To prove that $ts$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ we prove that every element of t is larger than every element of $s$. If $ts'$ is an occurrence of $\pmotif[\firsti{\pmotif}{i+1}:]$ so every element of $t$ is larger than every element of $s'$. Moreover the maximal element of $s$ is smaller than  the maximal element of $s'$ so every element of $s$ is smaller than every element of $s'$ thus every element of $s$ is smaller than every element of $t$. We use a similar argument if $\factor{\pmotif}{i}$ a descent factor.
\qed
\end{proof}

\begin{corollary}
Let $\pmotif$ be a permutation where $\factor{\pmotif}{i}$ is an ascent (resp. descent) factor.
If $s \in \SET{\pmotif}{\ptext}{i}{j}$, such that $s[\first{\pmotif}{i-1}]$ is minimal (resp. maximal). There is a match of $\pmotif$ in $\ptext$ with element $\first{\pmotif}{i}$ matched to $\ptext[j]$ if and only if there exists $t$ such that $ts$ is a match of $\pmotif$ in $\ptext$. 
\end{corollary}

% This corollary allows us to way to choose a match. Indeed if one does not find a match such that a well chosen subsequence is a suffix of the motif then there is no match at all.

@@

Consider the following problem :
$$
\LM{\pmotif}{\ptext}{i}{j} =
\begin{cases}

@@

	\text{The minimal value of a subsequence} 
		& \text{If $\first{\sigma}{i}$ is  } \\
	\text{ which is an occurence of $\sigma[\firsti{\sigma}{i}:]$ in $\pi[j:]$} & \text{ a descent element } \\	
	\text{ and that maximize the minimal value} & \\
	\text{ and $\first{\sigma}{i}$ is matched with $\ptext[j]$} & \\
\end{cases}
$$

This problem can be solve by induction as follows:

\noindent\textbf{BASE :}
\begin{align*}
\LM{\pmotif}{\ptext}{1}{j} 
&=
\begin{cases}
		\min_{j<j'}\{\infty\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LIS(j,j',\ptext[j'])  $}\}&\text{ an ascent element } \\
		&\\
		\max_{j<j'}\{0\} \cup \{\ptext[j'] |\text{ $j'$ such that \indent } 
			& \text{If $\first{\sigma}{i}$ is}  \\
		\text{ $|\factor{\pmotif}{1}| \leq LDS(j,j',\ptext[j']) $}\}&\text{ a descent element }\\
\end{cases}
\\
\intertext{\textbf{STEP :}}
\LM{\pmotif}{\ptext}{i}{j} 
&=
\begin{cases}
	\min \{\infty\} \cup  \AF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is an ascent element}\\
	\max \{0\} \cup  \DF{\pmotif}{\ptext}{i}{j} &
	\text{If $\first{\sigma}{i}$ is  a descent element}\\
\end{cases}
\end{align*}
where $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ are the sets 
of elements such that if  
$\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$ or if $\ptext[k] \in \DF{\pmotif}{\ptext}{i}{j}$ 
then there exists a match of $\pmotif[\firsti{\pmotif}{i}:]$ in $\ptext[j:]$ and the element 
$\first{\pmotif}{i-1}$ is matched with element $\ptext[k]$. 
Given that one has the set $\SET{\pmotif}{\ptext}{i-1}{k}$, 
to decide if $\ptext[k] \in \AF{\pmotif}{\ptext}{i}{j}$, 
one needs to find an occurrence $t$ of $\factor{\pmotif}{i}$ in $\ptext[j:k-1]$ 
such there exists $s \in \SET{\pmotif}{\ptext}{i-1}{k}$ and $ts$ is a match of 
$\pmotif[\firsti{\pmotif}{i}:]$. 
Formally we define $\AF{\pmotif}{\ptext}{i}{j}$ and $\DF{\pmotif}{\ptext}{i}{j}$ as follows:
\begin{align*}
\AF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq 0$ and} 
\\
&\qquad 
\text{$|\factor{\pmotif}{i}| \leq LIS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$} 
\\
\DF{\pmotif}{\ptext}{i}{j}
&=
\text{$\{\ptext[j'] \;|\; j<j'$ and $\LM{\pmotif}{\ptext}{i-1}{j'} \neq \infty$ and}
\\
&\qquad
\text{$|\factor{\pmotif}{i}| \leq LDS_{\ptext}(j,j'-1,\first{\pmotif}{i-1})\}$}
\end{align*}
where $LIS_{\ptext}(i,j,k)$ (resp. $LDS_{\ptext}(i,j,k)$) is the longest increasing 
(resp. decreasing) sequence in $\ptext$ starting at $i$ and ending at $j$,
with every element of this sequence
smaller (resp. bigger) than $k$.
$LIS_{\ptext}$ and $LDS_{\ptext}$ can be computed in 
$O(|\ptext|^2\log(\log(|\ptext|)))$ (see \cite{Bespamyatnikh00enumeratinglongest}).

For the base case,
one is looking for a match for the first factor.
If the factor is an ascent (respectively descent) factor, then
one has to find an increasing (resp. decreasing) subsequence
in the text of same size or longest that
the size of the first factor.
To find the "safest" solution
one must assure that the right most element
of the sequence is minimal (resp. maximal).\\
For the induction, it follows the same idea
except that one must assure that
every element of the subsequence is larger (resp. smaller)
than the maximal (resp. minimal) element of the rest of the match which is given by the inductive problem of
the previous factor.

There exists a match of $\pmotif$ in $\ptext$ if and only if
there exists a $LM(n,i)\neq 0$ and $LM(n,i)\neq \infty$ for $1 \leq i \leq |\ptext|$,
with $n$ the number of factor in $\pmotif$.
Moreover the basic case can be computed in $O(n^2)$
and the induction on $O(kn^2)$ time and space.
\qed
\end{proof}

@@

This section is devoted to presenting a polynomial-time algorithm for deciding whether
a $(213,231)$-avoiding bivincular pattern occurs in a permutation.
We start by presenting a property of bivincular motifs avoiding both $213$ and $231$.

\begin{lemma}
\label{lemma:ascentDescentAscent}
If $\overline{\sigma[i]\sigma[j]}$ (thus $\sigma[j]=\sigma[i]+1$) and 
if $\sigma[i]$ is an ascent (resp. decent) element and $\sigma[i]+1$ is an ascent 
(resp. decent) element then 
$i<j$ ($j>i$) and for every $k$, $i<k<j$ ($j>k>i$), $\pi[k]$ is a descent 
(resp. ascent) element.    
\end{lemma}

\begin{proof}
Suppose that there exists $k$, $i<k<j$,
such that $\pmotif[k]$ is ascent. Ascent elements are increasing  so $\sigma[i]<\sigma[k]<\sigma[j]$ which is in contradiction that with $\sigma[j]=\sigma[i]+1$.
We use a similar argument if $\sigma[i]$ is a descent element
\qed
\end{proof}

@@

\begin{proof}
We consider the following problem :
Given a bivincular motif $\pbmotif$ with $\pmotif \in \Av{n_\pmotif}{231,213} $, and a text $\ptext \in \Av{n_\ptext}{231,213}$.
$$
\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}=
\begin{cases}
	true 	& \text{If $\pbmotif[i:]$ occurs in $\ptext[j:]$}\\
			& \text{with every element is in $[\lb,\ub]$}\\
			& \text{and element $\pmotif[i]$ occurs in $\ptext[j]$ }\\
	false 	& otherwise\\
\end{cases}
$$

$\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ is closed under induction.
It can be solved by means by the following relations:

\noindent\textbf{BASE :} \\
$$
\PM{\pbmotif}{\ptext}{\lb}{\ub}{n_\pmotif}{j}=
\begin{cases}
	true 	& \text{if $\ptext[j] \in [\lb,\ub ]$}\\
			& \text{and if ${\pmotif[n_\pmotif]}_\lrcorner$ then $j=n_\ptext$}\\
			& \text{and if ${\pmotif[n_\pmotif]}^\urcorner$ then $\ptext[j]=\ub=n_\pmotif$}\\
			& \text{and if  $^\ulcorner{\pmotif[n_\pmotif]}$ then $\ptext[j]=\lb=1$ } \\
			& \text{and if  $\overline{(\pmotif[n_\pmotif]-1)\pmotif[n_\pmotif] }$ then $\ptext[j]=\lb$ }  \\
			& \text{and if  $\overline{(\pmotif[n_\pmotif]\pmotif[n_\pmotif]-1)}$ then $\ptext[j]=\ub$}  \\

@@

\noindent\textbf{STEP :}

We consider 3 cases for the problem $\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j}$ :
\begin{itemize}
	\item If $\ptext[j] \notin [\lb,\ub]$ then :
	$$
	\PM{\pbmotif}{\ptext}{\lb}{\ub}{i}{j} = false
	$$

@@

			the inductive problems after $\PM{\pbmotif}{\ptext}{*}{*}{i}{*}$ have for lower bound $\ptext[j]+1$. $\PM{\pbmotif}{\ptext}{*}{*}{i'}{*}$ is either an inductive case or a base case, in either way  $\PM{\pbmotif}{\ptext}{*}{*}{i'}{*}$ is true if and only if the occurrence of $\sigma[i']$ is equal to lower bound which is equal to $\sigma[i']+1$ so the occurrence of $\pmotif[i]$ and $\pmotif[i']$ are consecutive in value.
		\end{itemize}
\end{itemize}
\qed
\end{proof}
